# Missing values


```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```



```{r, echo=FALSE}
library(ggplot2)
library(dplyr)
library(ggridges)
library(tidyverse)
library(grid)
library(data.table)
library(scales)
library(GGally)
library(patchwork)
library(d3r)
library(parcoords)
library(ggalluvial)
library(vcd)
library(grid)
library(Lock5withR)
library(GGally)
library(hexbin)
library(MASS)
library(stringr) 
library(boot)
library(agridat)
library(openintro)
library(ggplot2)
library(dplyr)
library(ggridges)
library(tidyverse)
library(grid)
set.seed(5702)
mycars <- mtcars
mycars[1:25, "gear"] <- NA
mycars[10:20, 3:5] <- NA
for (i in 1:10) mycars[sample(32,1), sample(11,1)] <- NA
plot_missing <- function(df,percent) {
  missing_patterns <- data.frame(is.na(df)) %>%
    group_by_all() %>%
    count(name = "count", sort = TRUE) %>%
    ungroup()
  fullcol<-which(apply(missing_patterns, 1, function(r) any(r==TRUE))==FALSE)
  has_fullcol=TRUE
  if(length(fullcol)==0) {
    has_fullcol=FALSE
  }
  first_data<-sapply(df, function(y) sum(length(which(is.na(y)==TRUE))))
  ord<-order(first_data,decreasing=TRUE)
  second_data<-missing_patterns[ord]
  first_data<-first_data[ord]
  third_data<-missing_patterns$count
  d1 <- gather(df, col_name, value) %>%
  group_by(col_name) %>%
  summarize(missing= mean(is.na(value))*100) %>%
  arrange(desc(missing))
  d2 <- third_data/sum(third_data)*100
  real2<-gather(second_data,col_name, value)
  real2$row_num <- seq.int(nrow(real2))
  real2$row_num<-real2$row_num%%length(third_data)
  real2$row_num[real2$row_num==0]<-length(third_data)
  real2$row_num<-as.factor(real2$row_num)
  real2$col_name<-as.factor(real2$col_name)
  real2$value[real2$row_num==fullcol]<-2
  real2$value=as.factor(real2$value)
  lab=sapply((rev(seq.int(length(third_data)))),toString)
  
  colors <- c("#cbcbcb","#b69fe6","#b3b3b3")
  if (!has_fullcol) {
    colors<-c("#cbcbcb","#b69fe6")
  }
  p2<-ggplot(real2, aes(x = fct_relevel(col_name, colnames(df)[ord]), y = fct_relevel(row_num,lab), fill=value)) +
    geom_tile(color = "white",show_guide = FALSE) +
    labs(x="variable",y="missing pattern") +
    scale_fill_manual(values=colors) +
    annotate("text", x = length(first_data)/2, y = length(third_data)-fullcol+1, label = "complete cases") + theme_grey()+theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))
  
  newdf<-data.frame(num=first_data, key=colnames(df)[ord])
  p1<-0
  if(percent) {
    p1<-d1 %>%
  ggplot() +
    geom_bar(aes(fct_relevel(col_name, colnames(df)[ord]),missing), stat = 'identity',fill="#689af5", alpha = 0.7) + scale_y_continuous(breaks = seq(0, 100, by=25),limits=c(0,100)) +
    labs(x='variable', y="% rows missing:", title='Missing value patterns') +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.x = element_text(angle=45, vjust=1, hjust=1)
  )
  } else {
  p1<-newdf %>%
    ggplot() +
    geom_bar(aes(fct_relevel(key, colnames(df)[ord]),num), stat = 'identity',fill="#689af5", alpha = 0.7) + 
    labs(x='variable', y="num rows missing:", title='Missing value patterns') +
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      axis.text.x = element_text(angle=45, vjust=1, hjust=1)
    )
}
  
  colors2<-c("0"="#689af5","1"="#0751d5")
  if (!has_fullcol) {
    colors2<-c("#689af5")
  }
  p3<-0
  if(percent) {
   newdf2<-data.frame(x=as.factor(seq(1,length(d2),1)),num=d2,val=rep(0,length(d2) ))
    newdf2$val[fullcol]<-1
    newdf2$val<-as.factor(newdf2$val)
    p3<-newdf2 %>%
  ggplot() +
    geom_bar(aes(reorder(x,rev(seq(1,length(d2),1))),y=num,fill=val),stat = 'identity',position='dodge',alpha=0.7,show.legend = FALSE) +
  labs(y="% rows")+
    scale_fill_manual("legend",values =c("0"="#689af5","1"="#0751d5")) + 
  scale_y_continuous(breaks = seq(0, 100, by=25),limits=c(0,100)) +
   coord_flip()+
 
    theme(axis.title.y=element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA)
    
  ) 
  } else {
     newdf2<-data.frame(x=as.factor(seq(1,length(third_data),1)),num=third_data,val=rep(0,length(third_data)))
  newdf2$val[fullcol]<-1
  newdf2$val<-as.factor(newdf2$val)
    
  p3<-newdf2 %>%
    ggplot() +
    geom_bar(aes(reorder(x,rev(seq(1,length(third_data),1))),num,fill=val),stat = 'identity',position='dodge',alpha=0.7,show.legend = FALSE) +
    labs(y="row count")+
    scale_fill_manual("legend",values =colors2) +
    coord_flip() +
    theme(
      axis.title.y=element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.border = element_rect(colour = "gray39", fill=NA),
    axis.text.x = element_text(angle=45, vjust=1, hjust=1)
  )
  }
  
  plt <- p1+plot_spacer()+p2+p3+plot_layout(width=c(5,1),height=c(1,5))
  
  return(plt)
}


```




```{r, echo=FALSE}
NYCcrime<-read.csv("C:/Users/wxc/Downloads/NYPD_Complaint_Data_Current__Year_To_Date_.csv",na.strings=c("","NA"))


```


```{r}
crime <- NYCcrime%>%
  mutate(ExactDate = as.Date(CMPLNT_FR_DT, format = "%m/%d/%Y" )) %>%
  filter(ExactDate >= as.Date('2021-01-01'))
```

```{r}
library(lubridate)
crime$OFNS_DESC <- as.factor(crime$OFNS_DESC)

crime <- crime %>%
  mutate(
    CrimeType=fct_recode(OFNS_DESC,
                     "THEFT CRIMES"="PETIT LARCENY",
                     "HARRASSMENT"="HARRASSMENT 2",
                     "ASSAULT"="ASSAULT 3 & RELATED OFFENSES",
                     "CRIMINAL MISCHIEF"="CRIMINAL MISCHIEF & RELATED OF",
                     "THEFT CRIMES"="GRAND LARCENY",
                     "ASSAULT"="FELONY ASSAULT",
                     "OFFENSES AGAINST PUBLIC SENSIBILITIES"="OFF. AGNST PUB ORD SENSBLTY &",
                     "OTHER"="MISCELLANEOUS PENAL LAW",
                     "THEFT CRIMES"="ROBBERY",
                     "THEFT CRIMES"="BURGLARY",
                     "THEFT CRIMES"="GRAND LARCENY OF MOTOR VEHICLE",
                     "DRUGS"="DANGEROUS DRUGS",
                     "TRAFFIC OFFENSES"="VEHICLE AND TRAFFIC LAWS",
                     "OFFENSES AGAINST PUBLIC ADMINI"="OFFENSES AGAINST PUBLIC ADMINI",
                     "OTHER"="DANGEROUS WEAPONS",
                     "SEX CRIMES"="SEX CRIMES",
                     "FRAUDS"="FORGERY",
                     "FRAUDS"="THEFT-FRAUD",
                     "FRAUDS"="FRAUDS",
                     "FRAUDS"="OFFENSES INVOLVING FRAUD",
                     "ALCOHOL RELATED CRIMES"="INTOXICATED & IMPAIRED DRIVING",
                     "TRESPASS"="CRIMINAL TRESPASS",
                     "THEFT CRIMES"="POSSESSION OF STOLEN PROPERTY",
                     "TRAFFIC OFFENSES"="UNAUTHORIZED USE OF A VEHICLE",
                     "ASSAULT"="OFFENSES AGAINST THE PERSON",
                     "THEFT CRIMES"="RAPE",
                     "OTHER"="ADMINISTRATIVE CODE",
                     "ARSON"="ARSON",
                     "THEFT CRIMES"="OTHER OFFENSES RELATED TO THEF",
                     "HOMICIDE"="MURDER & NON-NEGL. MANSLAUGHTER",
                     "OTHER"="NYS LAWS-UNCLASSIFIED FELONY",
                     "OTHER"="OTHER STATE LAWS (NON PENAL LA",
                     "THEFT CRIMES"="BURGLAR'S TOOLS",
                     "THEFT CRIMES"="PETIT LARCENY OF MOTOR VEHICLE",
                     "GAMBLING"="GAMBLING",
                     "KIDNAPPING"="KIDNAPPING & RELATED OFFENSES",
                     "THEFT CRIMES"="THEFT OF SERVICES",
                     "OTHER"="AGRICULTURE & MRKTS LAW-UNCLASSIFIED",
                     "ALCOHOL RELATED CRIMES"="ALCOHOLIC BEVERAGE CONTROL LAW",
                     "OTHER"="OFFENSES AGAINST PUBLIC SAFETY",
                     "FRAUDS"="FRAUDULENT ACCOSTING",
                     "OTHER"="ANTICIPATORY OFFENSES",
                     "SEX CRIMES"="PROSTITUTION & RELATED OFFENSES",
                     "OTHER"="DISORDERLY CONDUCT",
                     "SEX CRIMES"="FELONY SEX CRIMES",
                     "OTHER"="ENDAN WELFARE INCOMP",
                     "OTHER"="OTHER STATE LAWS",
                     "OTHER"="NYS LAWS-UNCLASSIFIED VIOLATION",
                     "OTHER"="CHILD ABANDONMENT/NON SUPPORT",
                     "KIDNAPPING"="KIDNAPPING",
                     "OTHER"="OFFENSES RELATED TO CHILDREN",
                     "OTHER"="JOSTLING",
                     "ALCOHOL RELATED CRIMES"="INTOXICATED/IMPAIRED DRIVING",
                     "HOMICIDE"="HOMICIDE-NEGLIGENT,UNCLASSIFIE",
                     "OTHER"="ESCAPE 3",
                     "OTHER"="OFFENSES AGAINST MARRIAGE UNCL",
                     "OTHER"="NEW YORK CITY HEALTH CODE",
                     "HOMICIDE"="HOMICIDE-NEGLIGENT-VEHICLE",
                     "OTHER"="DISRUPTION OF A RELIGIOUS SERV",
                     "GAMBLING"="LOITERING/GAMBLING (CARDS, DIC",
                     "OTHER"="OTHER STATE LAWS (NON PENAL LAW)"
                     )) %>%
  
  mutate(ExactTime = as.ITime(CMPLNT_FR_TM)) %>%
  mutate(day=wday(ExactDate,label=TRUE,abbr=TRUE),
         month=month(ExactDate,label=TRUE,abbr=TRUE)) %>%
  
  mutate(Time = case_when(
    CMPLNT_FR_TM >="00:00:00" & CMPLNT_FR_TM <"01:00:00"~0,
    CMPLNT_FR_TM >="01:00:00" & CMPLNT_FR_TM <"02:00:00"~1,
    CMPLNT_FR_TM >="02:00:00" & CMPLNT_FR_TM <"03:00:00"~2,
    CMPLNT_FR_TM >="03:00:00" & CMPLNT_FR_TM <"04:00:00"~3,
    CMPLNT_FR_TM >="04:00:00" & CMPLNT_FR_TM <"05:00:00"~4,
    CMPLNT_FR_TM >="05:00:00" & CMPLNT_FR_TM <"06:00:00"~5,
    CMPLNT_FR_TM >="06:00:00" & CMPLNT_FR_TM <"07:00:00"~6,
    CMPLNT_FR_TM >="07:00:00" & CMPLNT_FR_TM <"08:00:00"~7,
    CMPLNT_FR_TM >="08:00:00" & CMPLNT_FR_TM <"09:00:00"~8,
    CMPLNT_FR_TM >="09:00:00" & CMPLNT_FR_TM <"10:00:00"~9,
    CMPLNT_FR_TM >="10:00:00" & CMPLNT_FR_TM <"11:00:00"~10,
    CMPLNT_FR_TM >="11:00:00" & CMPLNT_FR_TM <"12:00:00"~11,
    CMPLNT_FR_TM >="12:00:00" & CMPLNT_FR_TM <"13:00:00"~12,
    CMPLNT_FR_TM >="13:00:00" & CMPLNT_FR_TM <"14:00:00"~13,
    CMPLNT_FR_TM >="14:00:00" & CMPLNT_FR_TM <"15:00:00"~14,
    CMPLNT_FR_TM >="15:00:00" & CMPLNT_FR_TM <"16:00:00"~15,
    CMPLNT_FR_TM >="16:00:00" & CMPLNT_FR_TM <"17:00:00"~16,
    CMPLNT_FR_TM >="17:00:00" & CMPLNT_FR_TM <"18:00:00"~17,
    CMPLNT_FR_TM >="18:00:00" & CMPLNT_FR_TM <"19:00:00"~18,
    CMPLNT_FR_TM >="19:00:00" & CMPLNT_FR_TM <"20:00:00"~19,
    CMPLNT_FR_TM >="20:00:00" & CMPLNT_FR_TM <"21:00:00"~20,
    CMPLNT_FR_TM >="21:00:00" & CMPLNT_FR_TM <"22:00:00"~21,
    CMPLNT_FR_TM >="22:00:00" & CMPLNT_FR_TM <"23:00:00"~22,
    CMPLNT_FR_TM >="23:00:00" & CMPLNT_FR_TM <"24:00:00" ~23,
    CMPLNT_FR_TM >="24:00:00" ~24))
```


```{r}
NYC_target<-crime[,c("ExactTime","ExactDate","CrimeType","BORO_NM","Longitude","Latitude")]
NYC_target<-NYC_target %>% 
  rename(
    Borough=BORO_NM
    
    )

```



Missing pattern Plot for NYC crime data, both count and percentage plots:

```{r, echo=FALSE,fig.width = 10,fig.height=8}
plot_missing(NYC_target,FALSE)
```

```{r, echo=FALSE,fig.width = 10,fig.height=8}
plot_missing(NYC_target,TRUE)
```


In the missing pattern plot, there are only two missing patterns:

1, missing Borough

2, missing Crime Category



We can see that most of the cases are complete in this dataset, so all rows have missing values.

It can be observed from the top graph that 800 rows are missing Borough, and very few rous are missing Crime Category. However given the vast amount of crimes we can see that they are barely visible in the percentage plot


Also from the right part of the graph, almost 100% of rows are complete cases which have no missing values.

Since there are too few missing cases for Category, there is no need to analyse pattern for it.

Since there are still a few hundred cases with missing Borough, we can do some analysis on them


```{r}
NYC_dd<-NYC_target %>%
  group_by(CrimeType) %>%
  summarize(count=n()) %>%
  arrange(count)
```


```{r}
NYC_Borough<-NYC_target[is.na(NYC_target$Borough),] %>%
  group_by(CrimeType) %>%
  summarize(count=n()) %>%
  arrange(count)
NYC_Borough$percent<-NYC_Borough$count/NYC_dd[match(NYC_Borough$CrimeType,NYC_dd$CrimeType),]$count
```



```{r}
NYC_Borough %>% 
  ggplot(aes(x = reorder(CrimeType,count), y = count)) +
  geom_bar(stat = "identity", fill = "purple") +
  labs(x ="Crime Type", y = "Number of Crimes Missing Borough", title = "Type Pattern For Missing Borough") + 
  coord_flip()


```

```{r}
NYC_Borough %>% 
  ggplot(aes(x = reorder(CrimeType,percent), y = percent)) +
  geom_bar(stat = "identity", fill = "purple") +
  labs(x ="Crime Type", y = "Percentage of Crimes Missing Borough", title = "Type Pattern For Missing Borough") + 
  coord_flip()


```

By the count plot, we can see that the major crime types missing borough are theft and homicide.

By the percentage plot, We can see that over 90% of Homicide Crimes are missing borough information. This might because of the transportation of body, or that no borough wants to be record these vicious incidents so they just left it blank.

So when doing analysis, we should avoid discussing the correlation between homicide and borough since most cases are missing this piece of information.


For other categories, given such small percentage of missing cases, we can just ignore the missing cases and do analysis as usual. 


