# Missing values


```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```



```{r, echo=FALSE}

library(patchwork)
library(d3r)
library(parcoords)
library(ggalluvial)
library(vcd)
library(grid)
library(Lock5withR)
library(GGally)
library(hexbin)
library(MASS)
library(stringr) 
library(boot)
library(agridat)
library(openintro)
library(ggplot2)
library(dplyr)
library(ggridges)
library(tidyverse)
library(grid)
set.seed(5702)
mycars <- mtcars
mycars[1:25, "gear"] <- NA
mycars[10:20, 3:5] <- NA
for (i in 1:10) mycars[sample(32,1), sample(11,1)] <- NA
plot_missing <- function(df,percent) {
  missing_patterns <- data.frame(is.na(df)) %>%
    group_by_all() %>%
    count(name = "count", sort = TRUE) %>%
    ungroup()
  fullcol<-which(apply(missing_patterns, 1, function(r) any(r==TRUE))==FALSE)
  has_fullcol=TRUE
  if(length(fullcol)==0) {
    has_fullcol=FALSE
  }
  first_data<-sapply(df, function(y) sum(length(which(is.na(y)==TRUE))))
  ord<-order(first_data,decreasing=TRUE)
  second_data<-missing_patterns[ord]
  first_data<-first_data[ord]
  third_data<-missing_patterns$count
  d1 <- gather(df, col_name, value) %>%
  group_by(col_name) %>%
  summarize(missing= mean(is.na(value))*100) %>%
  arrange(desc(missing))
  d2 <- third_data/sum(third_data)*100
  real2<-gather(second_data,col_name, value)
  real2$row_num <- seq.int(nrow(real2))
  real2$row_num<-real2$row_num%%length(third_data)
  real2$row_num[real2$row_num==0]<-length(third_data)
  real2$row_num<-as.factor(real2$row_num)
  real2$col_name<-as.factor(real2$col_name)
  real2$value[real2$row_num==fullcol]<-2
  real2$value=as.factor(real2$value)
  lab=sapply((rev(seq.int(length(third_data)))),toString)
  
  colors <- c("#cbcbcb","#b69fe6","#b3b3b3")
  if (!has_fullcol) {
    colors<-c("#cbcbcb","#b69fe6")
  }
  p2<-ggplot(real2, aes(x = fct_relevel(col_name, colnames(df)[ord]), y = fct_relevel(row_num,lab), fill=value)) +
    geom_tile(color = "white",show_guide = FALSE) +
    labs(x="variable",y="missing pattern") +
    scale_fill_manual(values=colors) +
    annotate("text", x = length(first_data)/2, y = length(third_data)-fullcol+1, label = "complete cases") + theme_grey()+theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))
  
  newdf<-data.frame(num=first_data, key=colnames(df)[ord])
  p1<-0
  if(percent) {
    p1<-d1 %>%
  ggplot() +
    geom_bar(aes(fct_relevel(col_name, colnames(df)[ord]),missing), stat = 'identity',fill="#689af5", alpha = 0.7) + scale_y_continuous(breaks = seq(0, 100, by=25),limits=c(0,100)) +
    labs(x='variable', y="% rows missing:", title='Missing value patterns') +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.x = element_text(angle=45, vjust=1, hjust=1)
  )
  } else {
  p1<-newdf %>%
    ggplot() +
    geom_bar(aes(fct_relevel(key, colnames(df)[ord]),num), stat = 'identity',fill="#689af5", alpha = 0.7) + 
    labs(x='variable', y="num rows missing:", title='Missing value patterns') +
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      axis.text.x = element_text(angle=45, vjust=1, hjust=1)
    )
}
  
  colors2<-c("0"="#689af5","1"="#0751d5")
  if (!has_fullcol) {
    colors2<-c("#689af5")
  }
  p3<-0
  if(percent) {
   newdf2<-data.frame(x=as.factor(seq(1,length(d2),1)),num=d2,val=rep(0,length(d2) ))
    newdf2$val[fullcol]<-1
    newdf2$val<-as.factor(newdf2$val)
    p3<-newdf2 %>%
  ggplot() +
    geom_bar(aes(reorder(x,rev(seq(1,length(d2),1))),y=num,fill=val),stat = 'identity',position='dodge',alpha=0.7,show.legend = FALSE) +
  labs(y="% rows")+
    scale_fill_manual("legend",values =c("0"="#689af5","1"="#0751d5")) + 
  scale_y_continuous(breaks = seq(0, 100, by=25),limits=c(0,100)) +
   coord_flip()+
 
    theme(axis.title.y=element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA)
    
  ) 
  } else {
     newdf2<-data.frame(x=as.factor(seq(1,length(third_data),1)),num=third_data,val=rep(0,length(third_data)))
  newdf2$val[fullcol]<-1
  newdf2$val<-as.factor(newdf2$val)
    
  p3<-newdf2 %>%
    ggplot() +
    geom_bar(aes(reorder(x,rev(seq(1,length(third_data),1))),num,fill=val),stat = 'identity',position='dodge',alpha=0.7,show.legend = FALSE) +
    labs(y="row count")+
    scale_fill_manual("legend",values =colors2) +
    coord_flip() +
    theme(
      axis.title.y=element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.border = element_rect(colour = "gray39", fill=NA),
    axis.text.x = element_text(angle=45, vjust=1, hjust=1)
  )
  }
  
  plt <- p1+plot_spacer()+p2+p3+plot_layout(width=c(5,1),height=c(1,5))
  
  return(plt)
}


```




```{r, echo=FALSE}
NYCcrime<-read.csv("C:/Users/wxc/Downloads/NYPD_Complaint_Data_Current__Year_To_Date_.csv",na.strings=c("","NA"))


```


```{r, echo=FALSE}
NYC_target<-NYCcrime[,c("CMPLNT_FR_DT","OFNS_DESC","CMPLNT_FR_TM","BORO_NM","Longitude","Latitude")]
NYC_target<-NYC_target %>% 
  rename(
    Time = CMPLNT_FR_DT,
    Category = OFNS_DESC,
    Exact_Time=CMPLNT_FR_TM,
    Borough=BORO_NM
    
    )
```




Missing pattern Plot for NYC crime data, both count and percentage plots:

```{r, echo=FALSE,fig.width = 10,fig.height=8}
plot_missing(NYC_target,FALSE)
```

```{r, echo=FALSE,fig.width = 10,fig.height=8}
plot_missing(NYC_target,TRUE)
```



In the missing pattern plot, there are only two missing patterns:

1, missing Borough

2, missing Crime Category



We can see that most of the cases are complete in this dataset, so all rows have missing values.

It can be observed from the top graph that 800 rows are missing Borough, and very few rous are missing Crime Category. However given the vast amount of crimes we can see that they are barely visible in the percentage plot

Also from the right part of the graph, almost 100% of rows are complete cases which have no missing values.

Given such small number of missing cases, we do not have to make any extra changes to the dataset. 